---
title: "Week_6_Practical"
author: "Loubna"
date: "11/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Brief:
Part 1:  explore some techniques for analysing patterns of point data in R.

Part 2: explore spatial autocorrelation using R or ArcGIS. analyse the patterns of Blue Plaques — linking the buildings of the present to people of the past

## Question
“For any given London Borough, are the Blue Plaques within that borough distributed randomly or do they exhibit some kind of dispersed or clustered pattern?”


## load in m libraries
includes point pattern analysis tools found in spatstat
```{r}
library(spatstat)
library(here)
library(sp)
library(rgeos)
library(maptools)
library(GISTools)
library(tmap)
library(sf)
library(geojson)
library(geojsonio)
library(tmaptools)
library(stringr)
library(tidyverse)
library(raster)
library(fpc)
library(dbscan)
library(ggplot2)
library(OpenStreetMap)
```

## Load in SPATIAL DATA

```{r}
#1 get the London Borough Boundaries
LondonBoroughs <- st_read(here::here("Data", "statistical-gis-boundaries-london", "ESRI", "London_Borough_Excluding_MHW.shp"))


#2 get the location of all Blue Plaques in the City
BluePlaques <- st_read("https://s3.eu-west-2.amazonaws.com/openplaques/open-plaques-london-2018-04-08.geojson")%>%
  st_transform(.,27700)
```
## Extract and summarize, plot  London Borough

```{r}
#pull out london from london boroughs using string detect function

BoroughMap <- LondonBoroughs %>%
  dplyr::filter(str_detect(GSS_CODE, "^E09"))%>%
  st_transform(., 27700)

summary(BoroughMap)

#Draw a thematic map quickly
qtm(BoroughMap)


```

## Overlaying both: plot the blue plaques in the city

```{r}
#setting up t map for plotting
tmap_mode("plot")

#1 boroughmap
#2 map no color, transparency is 0.5
tm_shape(BoroughMap)+
  tm_polygons(col= NA, alpha = 0.5) +
tm_shape(BluePlaques)+
  tm_dots(col = "blue")
  
  
```

## Clean plaque data
remove duplicates and select only those inside london-clip

note: default is intersects, but we could also use:
BluePlaquesSub <- BluePlaques[BoroughMap, , op = st_within], 

```{r}
#remove duplicates (same grid refence.causes issue later in the analysis )
#sf and tidyverse
#BluePlaques <- distinct(BluePlaques)

BluePlaques <- BluePlaques%>%
  distinct(geometry, .keep_all=T)


#Now just select the points inside London:clipping to borough map
BluePlaquesSub <- BluePlaques[BoroughMap,]


#remap to see if they have been removed
tmap_mode("plot")
tm_shape(BoroughMap) +
  tm_polygons(col = NA, alpha = 0.5) +
tm_shape(BluePlaquesSub) +
  tm_dots(col = "blue")

```

## Select study Area for the analysis
1. subset our SpatialPolygonsDataFrame to pull out a borough we are interested in. I’m going to choose Harrow


```{r}
Harrow <- BoroughMap %>%
  filter(., NAME=="Harrow")  #select row where col 'NAME' is Harrow

#Check to see that the correct borough has been pulled out
tm_shape(Harrow) +
  tm_polygons(col = NA, alpha = 0.5)
```

## clip our Blue Plaques: fall within harrow

```{r}
#select the points inside Harrow :clipping to Harrow Boundary
BluePlaquesSub <- BluePlaques[Harrow,]

#check that it's worked

#sets up tmap for plotting
tmap_mode("plot")

#maps both the harrow shp file overlayed with clipped blue plaque points
tm_shape(Harrow) +
  tm_polygons(col = NA, alpha = 0.5) +
tm_shape(BluePlaquesSub) +
  tm_dots(col = "blue")
```

### Begin analysis using spatstat
this package reqires u to put a window in you study area

1.create an observation window
2. change sf to sp object. from sf to sp..original spatial pattern. then sp to point pattern

```{r}
#as ownin is as window..its a bounding box
window <- as.owin(Harrow)
plot(window)

```

```{r}

#convert sf to sp as spatstat does not work directly with the sf. For point pattern analysis, we need to create a point pattern (ppp) object

#to sp
BluePlaquesSub<- BluePlaquesSub %>%
  as(., 'Spatial')

#ppp pulls out the x and y coordinates 
BluePlaquesSub.ppp <- ppp(x=BluePlaquesSub@coords[,1],
                          y=BluePlaquesSub@coords[,2],
                          window=window)

```

```{r}
BluePlaquesSub@coords[,1]
```

```{r}
#point added to the window in order to begin analysis

BluePlaquesSub.ppp %>%
  plot(.,pch=16,cex=0.5,          
       main="Blue Plaques Harrow")  #title
```

##Point Pattern Analysis

#Kernel Density Estimation
plots the density of your points (ppp objects) under a window called a ‘Kernel’

```{r}
#sigma value sets the diameter of the Kernel.
BluePlaquesSub.ppp %>%
  density(., sigma=500) %>%
  plot()


```

#Quadrat Analysis
knowing whether the distribution of points in our study area differs from CSR(complete spat randomness) based on the poisson distribution. most basic test is quadrat analysis

```{r}
#First plot the points
plot(BluePlaquesSub.ppp,
     pch=16,
     cex=0.5, 
     main="Blue Plaques in Harrow")

#now count the points in that fall in a 6 x 6
#grid overlaid across the windowBluePlaquesSub.ppp2<-BluePlaquesSub.ppp %>%
BluePlaquesSub.ppp %>%
  quadratcount(.,nx = 6, ny = 6)%>%
    plot(., add=T, col="red")

```

```{r}
#save results of the quatrat analysis in a table. run quadrat count 
Qcount <- BluePlaquesSub.ppp %>%
  quadratcount(.,nx = 6, ny = 6) %>%
  as.data.frame() %>%
  dplyr::count(Var1=Freq)%>%
  dplyr::rename(Freqquadratcount=n)

#Check the data type in the first column
Qcount %>% 
  summarise_all(class)

#the table shows us no. of pts in each quad and the frequency count

```

```{r}
#next we calculate our expected values using poisson formula
sums <- Qcount %>%
  #calculate the total blue plaques (Var * Freq)
  mutate(total = Var1 * Freqquadratcount) %>%
  dplyr::summarise(across(everything(), sum))%>%
  dplyr::select(-Var1) 

lambda<- Qcount%>%
  #calculate lambda
  mutate(total = Var1 * Freqquadratcount)%>%
  dplyr::summarise(across(everything(), sum)) %>%
  mutate(lambda=total/Freqquadratcount) %>%
  dplyr::select(lambda)%>%
  pull(lambda)



QCountTable <- Qcount %>%
  mutate(Pr=((lambda^Var1)*exp(-lambda))/factorial(Var1))%>%
  #now calculate the expected counts based on our total number of plaques
  #and save them to the table
  mutate(Expected= (round(Pr * sums$Freqquadratcount, 0)))



#Compare the frequency distributions of the observed and expected point patterns
plot(c(1,5),c(0,14), type="n",
xlab="Number of Blue Plaques (Red=Observed,Blue=Expected)", 
     ylab="Frequency of Occurances")

points(QCountTable$Freqquadratcount, 
       col="Red", 
       type="o", 
       lwd=3)
points(QCountTable$Expected, col="Blue", 
       type="o", 
       lwd=3)


#seems like they allign..spatial randomness, no distinct clustering. but to know for sure, we need a chi square test
```
```{r}
#chi square test to know if random or clustered
#If our p-value is > 0.05 then this indicates that we have CSR and there is no pattern in our points. If it is < 0.05, this indicates that we do have clustering in our points.

#if you also add it to consol,, ull see the p value is 0.198
teststats <- quadrat.test(BluePlaquesSub.ppp, nx = 6, ny = 6)

plot(BluePlaquesSub.ppp,pch=16,cex=0.5, main="Blue Plaques in Harrow")
plot(teststats, add=T, col = "red")

#conclusion: there is no pattern in our points. but quatrat test isnt the best for spatial data so we will further investigate using riplesys k

```

#Ripleys K
One way of getting around the limitations of quadrat analysis is to compare the observed distribution of points with the Poisson random model for a whole range of different distance radii.

spatstat package using the kest() function


Red is the theoretical value of K for each distance window (r) under a Poisson assumption of Complete Spatial Randomness. The Black line is the estimated values of K accounting for the effects of the edge of the study area.

From the graph, we can see that up until distances of around 1300 metres, Blue Plaques appear to be clustered in Harrow, however, at around 1500 m, the distribution appears random and then dispersed between about 1600 and 2100 metres

```{r}
K <- BluePlaquesSub.ppp %>%
  Kest(., correction="border") %>%
  plot()
```
#Density based spatial clustering DBSCAN

Quadrat and Ripley’s K analysis tell  us if we have spatial clusters present in our point data, but dont tell us WHERE clusters are occurring. To discover this we need to use alternative techniques. One popular technique is DBSCAN

We will now carry out a DBSCAN analysis of blue plaques in my borough to see if there are any clusters present.


results of the Ripley’s K analysis earlier shows we are getting clustering up to a radius of around 1200m, with the largest bulge in the graph at around 700m radius . Therefore, 700m is probably a good place to start and we will begin by searching for clusters of at least 4 points…

```{r}
#first check the coordinates of the Harrow spatial polygon:
st_geometry(BoroughMap)

#then extract the points from the spatial points data frame.here we dont need ppp as we did for quad and ripleys
BluePlaquesSubPoints <- BluePlaquesSub %>%
  coordinates(.)%>%
  as.data.frame()

#now run the dbscan analysis
db <- BluePlaquesSubPoints %>%
  fpc::dbscan(.,eps = 700, MinPts = 4)

#now plot the results
plot(db, BluePlaquesSubPoints, main = "DBSCAN Output", frame = F)
plot(BoroughMap$geometry, add=T)


```

```{r}
# use kNNdistplot() from the dbscan pacakge to find a suitable eps value based on the ‘knee’ in the plot
# k is no of nearest neighbours used, use min points

BluePlaquesSubPoints%>%
  dbscan::kNNdistplot(.,k=4)
```
## Produce better DBSCAN map using ggplot

first add the cluster info in to the dataframe

```{r}
db
db$cluster

# now add this cluster membership info back into our dataframe

BluePlaquesSubPoints2<- BluePlaquesSubPoints %>%
  mutate(dbcluster=db$cluster)

```
Next: create convex hull polygons to wrap around the points in our clusters.

```{r}
chulls <- BluePlaquesSubPoints2 %>%
  group_by(dbcluster) %>%
  dplyr::mutate(hull = 1:n(),
  hull = factor(hull, chull(coords.x1, coords.x2)))%>%
  arrange(hull)


#0 isn’t actually a cluster so drop it from the dataframe
chulls <- chulls %>%
  filter(dbcluster >=1)


```

Now create a ggplot2 object from our data

```{r}

dbplot <- ggplot(data=BluePlaquesSubPoints2, 
                 aes(coords.x1,coords.x2, colour=dbcluster, fill=dbcluster)) 
#add the points in
dbplot <- dbplot + geom_point()
#now the convex hulls
dbplot <- dbplot + geom_polygon(data = chulls, 
                                aes(coords.x1,coords.x2, group=dbcluster), 
                                alpha = 0.5) 
#now plot, setting the coordinates to scale correctly and as a black and white plot 
#(just for the hell of it)...
dbplot + theme_bw() + coord_equal()

```
add a basemap
```{r}
#First get the bbox in lat long for Harrow
HarrowWGSbb <- Harrow %>%
  st_transform(., 4326)%>%
  st_bbox()

#add map from open street map
basemap <- OpenStreetMap::openmap(c(51.5549876,-0.4040502),c(51.6405356,-0.2671315),
                         zoom=NULL,
                         "stamen-toner")


# convert the basemap to British National Grid
basemap_bng <- openproj(basemap, projection="+init=epsg:27700")

```
 
```{r}

#Now we can plot our fancy map with the clusters on:
autoplot.OpenStreetMap(basemap_bng)+ 
  geom_point(data=BluePlaquesSubPoints, 
             aes(coords.x1,coords.x2, 
                 colour=dbcluster, 
                 fill=dbcluster)) + 
  geom_polygon(data = chulls, 
               aes(coords.x1,coords.x2, 
                   group=dbcluster,
                   fill=dbcluster), 
               alpha = 0.5
```



